### 2015-05-20 Harris_2015/manuscript-materials/draft.Rmd
version bump, manuscript tweaks, final figure

This paper only scratches the surface of what Markov networks can do for
ecology. This family of models---particularly the Ising model for binary
networks---has been extremely well-studied in statistical physics for nearly a
century, and the models' properties, capabilities, and limits are well-
understood in a huge range of applications, from spatial modeling
[@gelfand_modelling_2005] to neuroscience [@schneidman_weak_2006] to models of
human behavior [@lee_statistical_2013; salakhutdinov_restricted_2007].
Modeling species interactions using the same framework would thus allow
ecologists to tap into an enormous set of existing discoveries and techniques
for dealing with indirect effects, stability, and alternative stable states.

>>>>

This paper only scratches the surface of what Markov networks can do for
ecology. This family of models---particularly the Ising model for binary
networks---has been extremely well-studied in statistical physics for nearly a
century, and the models' properties, capabilities, and limits are well-
understood in a huge range of applications, from spatial modeling
[@gelfand_modelling_2005] to neuroscience [@schneidman_weak_2006] to models of
human behavior [@lee_statistical_2013; @salakhutdinov_restricted_2007].
Modeling species interactions using the same framework would thus allow
ecologists to tap into an enormous set of existing discoveries and techniques
for dealing with indirect effects, stability, and alternative stable states.

### 2015-05-20 Harris_2015/manuscript-materials/draft.tex
version bump, manuscript tweaks, final figure

These results also have important implications for the continued use of
fixed-fixed null models in ecology. The small simulated landscapes
described by Figure 1 show that test statistics based on marginal
co-occurrence (such as C-scores) will not always have a straightforward
relationship with the underlying ecological processes. Moreover, the
larger communities analyzed in Figure 3 often fell so far outside the
null distribution\footnote{Nearly 20\% of the species pairs fell outside
  the 99.99994\% confidence intervals implied by their \(Z\) scores
  (\(|Z| > 5\)), and about 10\% had uncorrected p-values below R's
  default numerical precision of \(2 \times 10^{-16}\).} that it
probably makes more sense to reject whole model rather than to assign
blame for the discrepancy to any one species pair. On average, the
pairwise \(Z\) scores from the null model provided less information
about direct species interactions than correlation coefficients did.
Researchers using the null modeling approach may be able to predict
twice as much of the variance in species' ``true'' interaction strengths
using partial covariances from linear regression or triple them using a
Markov network.

>>>>

These results also have important implications for the continued use of
fixed-fixed null models in ecology. The small simulated landscapes
described by Figure 1 show that test statistics based on marginal
co-occurrence (such as C-scores) will not always have a straightforward
relationship with the underlying ecological processes. Moreover, the
larger communities analyzed in Figure 3 often fell so far outside the
null distribution\footnote{Nearly 20\% of the species pairs fell outside
  the 99.99994\% confidence intervals implied by their \(Z\) scores
  (\(|Z| > 5\)), and about 10\% had uncorrected p-values below R's
  default numerical precision of \(2 \times 10^{-16}\).} that it
probably makes more sense to reject whole model rather than to assign
blame for the discrepancy to any one species pair. On average, the
pairwise \(Z\) scores from the null model provided less information
about direct species interactions than correlation coefficients did.
Researchers using null modeling approaches may be able to predict twice
as much of the variance in species' ``true'' interaction strengths using
partial covariances from linear regression, or triple them using a
Markov network.
